Deep Neural Networks to Sparse Variational Gaussian Processes

### 2023-04-24 paper/main.tex
updated abstract for new paper

  The world is a complex dynamic environment where humans and animals leverage high-dimensional sensory data to successfully operate and
  adapt to changes in the environment.
  Building learning-based systems that can handle high-dimensional data whilst adapting to new data whilst expressing uncertainty is essential for their deployment in real-world applications, e.g. healthcare and robotics.
  However, modern deep learning methods cannot adapt to new data because they rely on gradient-based optimisation, which requires retraining with the addition of new data. In contrast, nonparametric approaches -- such as Gaussian processes -- can adapt to new data. However, they are limited to low-dimensional and small data sets. We present a method that enables any trained neural network (NN) to be converted to a sparse variational Gaussian process (SVGP).

  However, modern deep learning methods cannot adapt to new data because they rely on gradient-based optimisation, which requires retraining with the addition of new data. In contrast, nonparametric approaches -- such as Gaussian processes -- can adapt to new data. However, they are limited to low-dimensional and small data sets. We present a method that enables any trained neural network (NN) to be converted to a sparse variational Gaussian process (SVGP).
Importantly, our method
1) scales to high-dimensional and large data sets,
2) provides well-calibrated uncertainty estimates, and
3) enables fast adaptation to new data.
Essentially, our method reaps the benefits of nonparametric models without their well-known issues.
We demonstrate our state-of-the-art NN uncertainty quantification with experiments using UCI data sets and
demonstrate the effectiveness of our model's fast adaptation by showing that it can improve sample-efficiency in model-based reinforcement learning.
%
% The world is a dynamic environment that requires humans and animals to constantly adapt.
%   Building learning-based systems that can adapt to new data and express uncertainty is essential for their deployment in real-world applications, e.g. healthcare and robotics. However, modern deep learning methods cannot adapt to new data because they rely on gradient-based optimisation, which requires retraining with the addition of new data. In contrast, nonparametric approaches -- such as Gaussian processes -- can adapt to new data. However, they are limited to low-dimensional and small data sets. We present a method that enables any trained neural network (NN) to be converted to a sparse variational Gaussian process (SVGP). Importantly, our method 1) enables fast adaptation to new data, 2) provides well-calibrated uncertainty estimates, and 3) scales to high-dimensional and large data sets. Essentially, our method reaps the benefits of nonparametric models without their well-known issues. We demonstrate the effectiveness of our model's fast adaptation by showing that it can improve sample-efficiency in model-based reinforcement learning and we demonstrate our state-of-the-art NN uncertainty quantification with experiments using UCI data sets.

>>>>

  % The world is a complex dynamic environment where humans and animals leverage high-dimensional sensory data to successfully operate and
  % adapt to changes in the environment.
  Building learning-based systems that can handle high-dimensional and large data sets whilst expressing uncertainty is essential in many real-world applications, e.g. healthcare and robotics.
  Although Gaussian processes can learn nonlinear functions whilst providing principled uncertainty quantification, they are limited to low-dimensional and small data sets.
  We present a method that enables any trained neural network (NN) to be converted to a sparse variational Gaussian process (SVGP).
Importantly, our method scales to high-dimensional and large data sets whilst providing well-calibrated uncertainty estimates.
Further to this, our function-space posterior enables
any neural network to be 1) adapted online without retraining and 2) used as a functional regulariser in continual learning.
Essentially, our method reaps the benefits of function space without the well-known issues associated with nonparametric models.
% Essentially, our method reaps the benefits of nonparametric models without their well-known issues.
We demonstrate that our method scales to high-dimensional and large data sets such as XXXXX.
We then show that our method obtains state-of-the-art uncertainty quantification with experiments using UCI data sets.
Finally, we demonstrate the power of our NN's function-space posterior by showing
that 1) it can improve sample-efficiency in model-based reinforcement learning (via our model's online adaptation)
and 2) it can improve XXXX in continual learning.

% Further to this, our method enables neural networks to be adapted online via efficient updates

% Essentially, our method reaps the benefits of nonparametric models without their well-known issues.

% We further show the benefits
% 3) enables fast adaptation to new data.

%   In contrast, nonparametric approaches -- such as Gaussian processes -- can adapt to new data. However, they are limited to low-dimensional and small data sets. We present a method that enables any trained neural network (NN) to be converted to a sparse variational Gaussian process (SVGP).

%   However, modern deep learning methods cannot adapt to new data because they rely on gradient-based optimisation, which requires retraining with the addition of new data.

%   However, modern deep learning methods cannot adapt to new data because they rely on gradient-based optimisation, which requires retraining with the addition of new data. In contrast, nonparametric approaches -- such as Gaussian processes -- can adapt to new data. However, they are limited to low-dimensional and small data sets.
%   We present a method that enables any trained neural network (NN) to be converted to a sparse variational Gaussian process (SVGP).
% Importantly, our method
% 1) scales to high-dimensional and large data sets,
% 2) provides well-calibrated uncertainty estimates, and
% 3) enables fast adaptation to new data.
% Essentially, our method reaps the benefits of nonparametric models without their well-known issues.
% We demonstrate our state-of-the-art NN uncertainty quantification with experiments using UCI data sets and
% demonstrate the effectiveness of our model's fast adaptation by showing that it can improve sample-efficiency in model-based reinforcement learning.
%
% The world is a dynamic environment that requires humans and animals to constantly adapt.
%   Building learning-based systems that can adapt to new data and express uncertainty is essential for their deployment in real-world applications, e.g. healthcare and robotics. However, modern deep learning methods cannot adapt to new data because they rely on gradient-based optimisation, which requires retraining with the addition of new data. In contrast, nonparametric approaches -- such as Gaussian processes -- can adapt to new data. However, they are limited to low-dimensional and small data sets. We present a method that enables any trained neural network (NN) to be converted to a sparse variational Gaussian process (SVGP). Importantly, our method 1) enables fast adaptation to new data, 2) provides well-calibrated uncertainty estimates, and 3) scales to high-dimensional and large data sets. Essentially, our method reaps the benefits of nonparametric models without their well-known issues. We demonstrate the effectiveness of our model's fast adaptation by showing that it can improve sample-efficiency in model-based reinforcement learning and we demonstrate our state-of-the-art NN uncertainty quantification with experiments using UCI data sets.
